Probabilistic Inference of NYC Congestion from Taxi Data

### 2017-12-10 reports/paper/paper.tex
intros

% re

\section{Background}
% {
% \color{gray}
% Example Structure:
% \begin{itemize}
% \item What information does a non-expert need to know about the problem domain?
% \item What data exists for this problem?
% \item What are the challenges/opportunities inherent to the data? (High dimensional, sparse, missing data, noise, structure, discrete/continuous, etc?)
% \end{itemize}
% }

@@

\section{Related Work}

@@

\section{Model}

@@

\section{Methods}

@@

\section{Results}

@@

\section{Discussion}

@@

\section{Conclusion}

>>>>

Estimating congestion and predicting trip time are important problems in urban economics, civil engineering, and operations research. The problem lends well to approaches in computational probabilistic inference and machine learning, thanks to the increasing availability of GPS data. Non-probabilistic machine learning methods, such as neural networks, perform extremely well in prediction, but generate parameters that do not have structural interpretations. On the other hand, current probabilistic methods, most of which are based on mixture models, either rely on extremely granular data such as closely spaced GPS measurements \cite{hunter2009path} or are difficult to scale to a large geographical area \cite{zhan2013urban}.

Our main contribution in this project is devising a stylized probabilistic model of travel, congestion, and route choice and creating two strategies of parameterization and approximate inference. From a data standpoint, our model does not require extremely granular data---training on a large dataset of taxi rides in New York City that only includes start- and end-trip locations. From an implementation standpoint, our model is parsimonious, scalable to large geographical areas and large datasets, and easy to train. From a prediction standpoint, our model is close to or outperforms many baseline metrics, such as neural network, linear regression, or Google Maps. In stark contrast with black-box methods, our model generates interpretable parameters that directly corresponds to level of congestion in a small area. We believe the ability to generate easily interpretable parameters is the foremost of our contributions, as it is transparent and accessible to a broad class of users---leading to, for instance, better urban planning and transportation design by policymakers.

This report proceeds as follows. Sections~\ref{sec:backgrd} and \ref{sec:related} briefly discuss some features of the data and some work in the previous literature. Section~\ref{sec:model} introduces our modeling framework formally and Section~\ref{sec:inference} introduces two methods for approximate inference of our model. Sections~\ref{sec:methods} and \ref{sec:results} discuss training on the TLC dataset and present predictive metrics of the training results relative to baseline. Section~\ref{sec:discuss} discusses the results and the accompanying visualization of parameters, while addressing issues of concern and of future inquiry. Section~\ref{sec:conc} concludes.

\section{Background}
\label{sec:backgrd}
% {
% \color{gray}
% Example Structure:
% \begin{itemize}
% \item What information does a non-expert need to know about the problem domain?
% \item What data exists for this problem?
% \item What are the challenges/opportunities inherent to the data? (High dimensional, sparse, missing data, noise, structure, discrete/continuous, etc?)
% \end{itemize}
% }

@@

\section{Related Work}
\label{sec:related}

@@

\section{Model}
\label{sec:model}

@@

\section{Methods}
\label{sec:methods}

@@

\section{Results}
\label{sec:results}

@@

\section{Discussion}
\label{sec:discuss}

@@

\section{Conclusion}
\label{sec:conc}
